Minfi Workflow
==============================
#Getting Start
```
library(minfi)
```
#Read Processed Data
```{r}
mset<-getGenomicRatioSetFromGEO('GSE42865')
class(mset)
```

#Read signal intensities data
```{r}
path <- getGEOSuppFiles('GSE42865')
print(path)
```

```{r}
filepath <- row.names(path)[2]
unzipped_filepath <- gsub("[.]gz$", "", filepath)
gunzip(filepath,destname=unzipped_filepath)
gmset=readGEORawFile(unzipped_filepath,Uname="Signal_A",Mname="Signal_B",sep="\t")
class(gmset)
```

Merge `pData` into raw intensity data
```{r}
pd <- pData(mset)
m <- match(rownames(pData(gmset)), rownames(pd))
pData(gmset) <- pd[m,]
```

#What is the difference?
Basically, they are the same. Both can be applied to most of the functions listed in this reporte preprocessMethod() can tell us what method is applied for both data (it is useful for GRChannelSet, too)


Difference: 
  1. the value is slightly different, even though the preprocessed method is the same.  
  2. data from signal intensities data do not contain good pData.  
  3. In their matrix, the order of samples and probes is different.  

#Preprocessing
We can check what kind of preprocess method is applied for the data.
```{r}
preprocessMethod(mset)
preprocessMethod(gmset)
```
There is no preprocess method suitable for preprocessed data, and it do not need preprocessed, too.
Only preprocessQuantile() can be applied to signal intensities data.
```{r}
qgmset=preprocessQuantile(gmset)
```

#Phenotype data
DESCRIBE THE DATA HERE

```{r}
pd <- pData(qgmset)
cols <- grep("characteristics_ch1", colnames(pd))
tab <- pd[,cols]

extract <- function(x) {
  tmp <- strsplit(as.character(x), split=": ")
  colname <- tmp[[1]][1]
  values <- sapply(tmp,"[",2)
  list(colname, values)
}
tmp <- lapply(tab, extract)
tab2 <- as.data.frame(sapply(tmp, "[[", 2))
names(tab2) <- sapply(tmp, "[[", 1)
pData(qgmset) <- DataFrame(cbind(pd, tab2))
```

```{r}
pd <- pData(qgmset)
table(pd$gender)
table(pd$health, pd$cell)
```

#Data visualization(Multi-dimension scaling plots)
```{r}
mdsPlot(getM(qgmset),numPositions=1000,sampGroups=pd$health,sampNames=pd$health)
```

#Data visualization(Sample Distance (Euclidean))
```{r}
sampleDists<-dist(t(getBeta(mset)))  
library(gplots)
library('RColorBrewer')
sampleDistMatrix<-as.matrix(sampleDists)
colors<-colorRampPalette(rev(brewer.pal(9,'Blues')))(255)
hc<-hclust(sampleDists)
heatmap.2(sampleDistMatrix,Rowv=as.dendrogram(hc),symm=TRUE,trace='none',col=colors,margins=c(2,10),labCol=FALSE)
```

#Differentially Methylated Position
```{r}
samplesToKeep <- grepl("EBV", pd$cell)
gmset2 <- qgmset[,samplesToKeep]
pData(gmset2)$status <- ifelse(pData(gmset2)$health == "Healthy donor", "control", 
                               ifelse(pData(gmset2)$health == "Werner Syndrome", "werner", "progeria"))
dmp<-dmpFinder(getM(gmset2)[,pData(gmset2)$status != "werner"], 
               pheno=pData(gmset2)$status[pData(gmset2)$status != "werner"],
               type='categorical')
```

#Find the Block
Only data from signal intensities data can be applied
```{r}
shortcpg<-cpgCollapse(gmset[,1:4],what='Beta',returnBlockInfo=FALSE)
design=matrix(c(1,0,1,0,0,1,0,1),nrow=4,ncol=2,byrow=TRUE)
rownames(design)=c(1,2,3,4)
colnames(design)=c('A','B')
block=blockFinder(shortcpg,design=design,what='Beta',cutoff=0.0001,smooth=FALSE)
```

#Diagnostic Plots(Plot the genes)
```{r}
cpgs <- rownames(dmp)[1]
plotCpg(getM(mset[,1:4]), cpg=cpgs, pheno=c('A','A','B','B'),measure='M')
```

#Diagnostic Plots(MAplot)
```{r}
library(affy)
rowMeans(log2(getBeta(mset[1:500,1:2])))->A
M<-log2(getBeta(mset[1:500,1])/getBeta(mset[1:500,2]))
intersect(which(M!='NaN'),which(A!='NaN'))->filter
intersect(filter,which(M!='Inf'))->filter
intersect(filter,which(M!='-Inf'))->filter
intersect(filter,which(A!='-Inf'))->filter
intersect(filter,which(A!='Inf'))->filter
ma.plot(A[filter],M[filter],cex=1)
```

#Diagnostic Plots(histogram)
```{r}
hist(dmp$qval,breaks=20,col="grey50", border="white")
```

#Gene Clustering
```{r}
library(RColorBrewer)
library("genefilter")
library(gplots)
sidecols<-c('grey','grey','dodgerblue','dodgerblue')
topVarGenes <-head(rownames(dmp),30)
colors<-colorRampPalette(rev(brewer.pal(9,'PuOr')))(255)
mat<-getBeta(mset)[topVarGenes,1:4]
mat<-mat-rowMeans(mat)
heatmap.2(mat,trace='none',col=colors,ColSideColors=sidecols,labRow=FALSE,mar=c(10,2),scale='row')
```


#Plotting in genomic space
```{r}
library(Gviz)
cpgIslands<-getLocations(mset[1:5,]) # To get the range information
chr <- as.character(unique(seqnames(cpgIslands)))
gen <- genome(cpgIslands)
atrack <- AnnotationTrack(cpgIslands, name = "CpG")            
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr)                     
plotTracks(list(itrack, gtrack, atrack))
```

#Epiviz
Aim for [this](http://epiviz.cbcb.umd.edu/?ws=WZCPGTd7Duv&seqName=chr11&start=130099272&end=133637378&settings=default&)